---
title: "Music Analysis Dashboard"
author: Alex
output: 
  flexdashboard::flex_dashboard:
    theme: 
      version: 4
      bootswatch: litera
      orientation: columns
      vertical_layout: fill
---

```{r setup, include=FALSE}
library(tidyverse)
library(flexdashboard)
library(ggplot2)
library(plotly)
library(rjson)
library(tidymodels)
library(ggdendro)
library(heatmaply)
library(sparsevctrs)
source("compmus.R")

major_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    0,    0)
minor_chord <-
  c(   1,    0,    0,    1,    0,    0,    0,    1,    0,    0,    0,    0)
seventh_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    1,    0)

major_key <-
  c(5.0, 2.0, 3.5, 2.0, 4.5, 4.0, 2.0, 4.5, 2.0, 3.5, 1.5, 4.0)
minor_key <-
  c(5.0, 2.0, 3.5, 4.5, 2.0, 4.0, 2.0, 4.5, 3.5, 2.0, 1.5, 4.0)

chord_templates <-
  tribble(
    ~name, ~template,
    "Gb:7", circshift(seventh_chord, 6),
    "Gb:maj", circshift(major_chord, 6),
    "Bb:min", circshift(minor_chord, 10),
    "Db:maj", circshift(major_chord, 1),
    "F:min", circshift(minor_chord, 5),
    "Ab:7", circshift(seventh_chord, 8),
    "Ab:maj", circshift(major_chord, 8),
    "C:min", circshift(minor_chord, 0),
    "Eb:7", circshift(seventh_chord, 3),
    "Eb:maj", circshift(major_chord, 3),
    "G:min", circshift(minor_chord, 7),
    "Bb:7", circshift(seventh_chord, 10),
    "Bb:maj", circshift(major_chord, 10),
    "D:min", circshift(minor_chord, 2),
    "F:7", circshift(seventh_chord, 5),
    "F:maj", circshift(major_chord, 5),
    "A:min", circshift(minor_chord, 9),
    "C:7", circshift(seventh_chord, 0),
    "C:maj", circshift(major_chord, 0),
    "E:min", circshift(minor_chord, 4),
    "G:7", circshift(seventh_chord, 7),
    "G:maj", circshift(major_chord, 7),
    "B:min", circshift(minor_chord, 11),
    "D:7", circshift(seventh_chord, 2),
    "D:maj", circshift(major_chord, 2),
    "F#:min", circshift(minor_chord, 6),
    "A:7", circshift(seventh_chord, 9),
    "A:maj", circshift(major_chord, 9),
    "C#:min", circshift(minor_chord, 1),
    "E:7", circshift(seventh_chord, 4),
    "E:maj", circshift(major_chord, 4),
    "G#:min", circshift(minor_chord, 8),
    "B:7", circshift(seventh_chord, 11),
    "B:maj", circshift(major_chord, 11),
    "D#:min", circshift(minor_chord, 3)
  )

key_templates <-
  tribble(
    ~name, ~template,
    "Gb:maj", circshift(major_key, 6),
    "Bb:min", circshift(minor_key, 10),
    "Db:maj", circshift(major_key, 1),
    "F:min", circshift(minor_key, 5),
    "Ab:maj", circshift(major_key, 8),
    "C:min", circshift(minor_key, 0),
    "Eb:maj", circshift(major_key, 3),
    "G:min", circshift(minor_key, 7),
    "Bb:maj", circshift(major_key, 10),
    "D:min", circshift(minor_key, 2),
    "F:maj", circshift(major_key, 5),
    "A:min", circshift(minor_key, 9),
    "C:maj", circshift(major_key, 0),
    "E:min", circshift(minor_key, 4),
    "G:maj", circshift(major_key, 7),
    "B:min", circshift(minor_key, 11),
    "D:maj", circshift(major_key, 2),
    "F#:min", circshift(minor_key, 6),
    "A:maj", circshift(major_key, 9),
    "C#:min", circshift(minor_key, 1),
    "E:maj", circshift(major_key, 4),
    "G#:min", circshift(minor_key, 8),
    "B:maj", circshift(major_key, 11),
    "D#:min", circshift(minor_key, 3)
  )

source("compmus.R")

get_conf_mat <- function(fit) {
  outcome <- .get_tune_outcome_names(fit)
  fit |> 
    collect_predictions() |> 
    conf_mat(truth = outcome, estimate = .pred_class)
}  

get_pr <- function(fit) {
  fit |> 
    conf_mat_resampled() |> 
    group_by(Prediction) |> mutate(precision = Freq / sum(Freq)) |> 
    group_by(Truth) |> mutate(recall = Freq / sum(Freq)) |> 
    ungroup() |> filter(Prediction == Truth) |> 
    select(class = Prediction, precision, recall)
}  
```


Introduction - Class Corpus {.storyboard}
=========================================
### Introducing my own tracks

I will structure my portfolio this week, but for now please go to the tempogram section in the menu. Thanks!

### Class Corpus Visualization

```{r}
library(readr)
library(ggplot2)
library(plotly)
library(viridis)

compmus2025 <- read_csv("C:\\Users\\aneru\\musika2\\compmus2025.csv")

# Build the ggplot object
p <- compmus2025 |>
  ggplot(
    aes(
      x = valence,
      y = arousal,
      colour = danceability
    )
  ) +
  geom_point(size = 2.5, alpha = 0.8) +  # Adjust size & transparency
  geom_smooth(method = "loess", se = FALSE, linewidth = 1, colour = "gray30") +  # Clean smooth
  geom_rug(linewidth = 0.2, alpha = 0.5) +  # Add data distribution markers
  scale_x_continuous(
    limits = c(3, 5),
    breaks = c(3, 4, 5, 6, 7),
    minor_breaks = NULL
  ) +
  scale_y_continuous(
    limits = c(3, 6),
    breaks = c(3, 4, 5, 6, 7, 8),
    minor_breaks = NULL
  ) +
  scale_colour_viridis_c(option = "turbo") +  # More colorful variant
  theme_light(base_size = 14) +
  labs(
    title = "Valence vs. Arousal",
    subtitle = "Colored by Danceability",
    x = "Valence",
    y = "Arousal",
    colour = "Danceability"
  ) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  )

# Make it interactive
ggplotly(p)

```
------------------------------------------------------------------------

This graph shows the relationship between Valence (how happy or positive a song sounds) and Arousal (how energetic a song is). The x-axis represents Valence, and the y-axis represents Arousal.

Each dot on the graph represents a song. The color of the dots shows Danceability, with purple meaning less danceable and*yellow meaning more danceable. Some dots are bigger than others, which could mean those songs are more popular or appear more often.

There is a blue trend line in the graph, showing that as Valence increases (songs sound happier), Arousal also tends to increase (songs have more energy). The gray area around the line shows some uncertainty, meaning not every song follows this trend exactly.

On the sides of the graph, there are small tick marks (rug plots) that show how many songs fall into different Valence and Arousal levels. 

Overall, the graph suggests that happier songs tend to be more energetic, and danceability varies, with some of the most danceable songs appearing in the high-energy, happy range.

### Clustering

```{r}
compmus2025 <- read_csv("C:\\Users\\aneru\\musika2\\compmus2025.csv")
cluster_juice <-
  recipe(
    filename ~
      arousal +
      danceability +
      instrumentalness +
      tempo +
      valence,
    data = compmus2025
  ) |>
  step_center(all_predictors()) |>
  step_scale(all_predictors()) |> 
  # step_range(all_predictors()) |> 
  prep(compmus2025) |>
  juice() |>
  column_to_rownames("filename")
compmus_dist <- dist(cluster_juice, method = "euclidean")
compmus_dist |> 
  hclust(method = "single") |> # Try single, average, and complete.
  dendro_data() |>
  ggdendrogram()
```
### Full Class Corpus Heatmap

```{r}
# Show every 8th row name only
row_labels <- rownames(cluster_juice)
custom_row_labels <- ifelse(seq_along(row_labels) %% 8 == 1, row_labels, "")

# Clean and readable heatmap
heatmaply(
  cluster_juice,
  dist_method = "euclidean",
  hclust_method = "average",
  k_col = 4,
  k_row = 4,
  plot_method = "plotly"
)
```

### My Tracks in Classcorpus Heatmap
```{r}
# Ensure reproducibility
set.seed(123)

# Define rows to include
highlight_names <- c("aleksandra-b-1", "aleksandra-b-2")
all_rows <- rownames(cluster_juice)

# Filter to keep those two
mandatory_rows <- all_rows[tolower(all_rows) %in% tolower(highlight_names)]

# Sample 8 more rows randomly (excluding the highlighted ones)
remaining_rows <- setdiff(all_rows, mandatory_rows)
random_rows <- sample(remaining_rows, size = 8)

# Combine into final subset
selected_rows <- c(mandatory_rows, random_rows)
subset_data <- cluster_juice[selected_rows, ]

# Plot the heatmap
heatmaply(
  subset_data,
  dist_method = "euclidean",
  hclust_method = "average",
  k_col = 2,  # fewer clusters since we have fewer rows
  k_row = 2,
  plot_method = "plotly",
  row_text_angle = 0,
  column_text_angle = 45,
  fontsize_row = 12,
  fontsize_col = 12,
  margins = c(80, 150)
)

```

```{r}
compmus2025_filtered <- 
  compmus2025 |> filter(!is.na(ai)) |> 
  mutate(ai = factor(if_else(ai, "AI", "Non-AI")))
```
```{r}
classification_recipe <-
  recipe(
    ai ~
      arousal +
      danceability +
      instrumentalness +
      tempo +
      valence,
    data = compmus2025_filtered
  ) |>
  step_center(all_predictors()) |>
  step_scale(all_predictors())      # Converts to z-scores.
  # step_range(all_predictors())    # Sets range to [0, 1].
```
```{r}
compmus_cv <- compmus2025_filtered |> vfold_cv(5)
```

```{r}
knn_model <-
  nearest_neighbor(neighbors = 1) |>
  set_mode("classification") |> 
  set_engine("kknn")
classification_knn <- 
  workflow() |> 
  add_recipe(classification_recipe) |> 
  add_model(knn_model) |> 
  fit_resamples(compmus_cv, control = control_resamples(save_pred = TRUE))
```

### k-Nearest Neighbour
```{r}
classification_knn |>
  get_conf_mat() |>
  autoplot(type = "heatmap") +
  scale_fill_gradient2(
    low = "#00bfc4",  # Teal
    mid = "white",
    high = "#f8766d", # Orange-red
    midpoint = 25,
    name = "Count"
  ) +
  labs(
    title = "Confusion Matrix - Custom Colors",
    x = "Predicted Class",
    y = "True Class"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    panel.grid = element_blank()
  )

```
### Random Forests

```{r}
forest_model <-
  rand_forest() |>
  set_mode("classification") |> 
  set_engine("ranger", importance = "impurity")
indie_forest <- 
  workflow() |> 
  add_recipe(classification_recipe) |> 
  add_model(forest_model) |> 
  fit_resamples(
    compmus_cv, 
    control = control_resamples(save_pred = TRUE)
  )
indie_forest |> get_pr()
```
```{r}
compmus2025_filtered |>
  ggplot(aes(x = valence, y = arousal, colour = ai, size = tempo)) +
  geom_point(alpha = 0.8) +
  scale_color_viridis_d() +
  labs(
    x = "Valence",
    y = "Arousal",
    size = "Tempo",
    colour = "AI"
  )
```

Chromograms and cepstpgrams {.storyboard}
=========================================
### Chromagram of my first track
```{r fig.show="hold", out.width="49%", fig.align="center"}
"features/aleksandra-b-1.json" |>                       # Changed the track to mine
  compmus_chroma(norm = "manhattan") |>                 # Change the norm
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
   scale_x_continuous(
    breaks = seq(0, 200, 50),  # More structured x-axis ticks
    minor_breaks = NULL
  ) +
  scale_fill_viridis_c(option = "inferno", begin = 0.2, end = 0.8) +  # More readable color scale
  labs(
    title = "Chromagram of my first track",
    x = "Time (s)",
    y = "Pitch Class",
    fill = "Intensity"
  ) +
  theme_minimal(base_size = 14) +  
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    panel.grid.major = element_line(color = "gray80", size = 0.3),
    panel.grid.minor = element_blank()
  )
```

### Chromagram of my second track
```{r fig.show="hold", out.width="49%", fig.align="center"}
"features/aleksandra-b-2.json" |>                       # Changed the track to mine
  compmus_chroma(norm = "manhattan") |>                 # Change the norm
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
   scale_x_continuous(
    breaks = seq(0, 200, 50),  # More structured x-axis ticks
    minor_breaks = NULL
  ) +
  scale_fill_viridis_c(option = "inferno", begin = 0.2, end = 0.8) +  # More readable color scale
  labs(
    title = "Chromagram of my second track",
    x = "Time (s)",
    y = "Pitch Class",
    fill = "Intensity"
  ) +
  theme_minimal(base_size = 14) +  
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    panel.grid.major = element_line(color = "gray80", size = 0.3),
    panel.grid.minor = element_blank()
  )
```

### Cepstogram of my first track
```{r}
"features/aleksandra-b-1.json" |>                       # Track file
  compmus_mfccs(norm = "identity") |>                  # MFCC computation
  ggplot(aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = seq(0, 12, 2),  # More structured y-axis
    minor_breaks = NULL
  ) +
  scale_x_continuous(
    breaks = seq(0, 200, 50), # More structured x-axis
    minor_breaks = NULL
  ) +
  scale_fill_distiller(palette = "Spectral", direction = 1) + # Better color contrast
  labs(
    title = "Cepstrogram of my first track",
    x = "Time (s)",
    y = "Coefficient Number",
    fill = "Intensity"
  ) +
  theme_minimal(base_size = 14) +  
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    panel.grid.major = element_line(color = "gray80", size = 0.3),
    panel.grid.minor = element_blank()
  )          
```

### Cepstogram of my second track
```{r}

"features/aleksandra-b-2.json" |>                       # Track file
  compmus_mfccs(norm = "identity") |>                  # MFCC computation
  ggplot(aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = seq(0, 12, 2),  # More structured y-axis
    minor_breaks = NULL
  ) +
  scale_x_continuous(
    breaks = seq(0, 200, 50), # More structured x-axis
    minor_breaks = NULL
  ) +
  scale_fill_distiller(palette = "Spectral", direction = 1) + # Better color contrast
  labs(
    title = "Cepstrogram of my second track",
    x = "Time (s)",
    y = "Coefficient Number",
    fill = "Intensity"
  ) +
  theme_minimal(base_size = 14) +  
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    panel.grid.major = element_line(color = "gray80", size = 0.3),
    panel.grid.minor = element_blank()
  ) 
```


Self-Similarity Matrices {.storyboard}
=========================================
### Chroma-based self-similarity for my first track

```{r, fig.width=14, fig.height=7}
"features/aleksandra-b-1.json" |>                     # Change the track
  compmus_chroma(norm = "identity") |>                # Change the norm
  compmus_self_similarity(
    feature = pc,
    distance = "euclidean"                            # Change the distance metric
  ) |>
  
  # Use ggplot to visualize the matrix
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_tile() +                                       # Better visualization than geom_raster()
  
  # Choose a better color scale
  scale_fill_viridis_c(guide = "none") +

  # Improve axis labels
  scale_x_continuous(
    breaks = seq(0, 200, 50),
    minor_breaks = NULL
  ) +
  scale_y_continuous(
    breaks = seq(0, 200, 50),
    minor_breaks = NULL
  ) +

  # Improve labels and theme
  labs(
    title = "Self-Similarity Matrix of My First Track",
    x = "Time (s)",
    y = NULL,
    fill = "Distance"
  ) +
  theme_minimal(base_size = 14) +  
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    panel.grid.major = element_line(color = "gray80", size = 0.3),
    panel.grid.minor = element_blank()
  )    
```

### Chroma-based self-similarity for my second track

```{r, fig.width=14, fig.height=7}
"features/aleksandra-b-2.json" |>                     # Change the track
  compmus_chroma(norm = "identity") |>                # Change the norm
  compmus_self_similarity(
    feature = pc,
    distance = "euclidean"                            # Change the distance metric
  ) |>
  
  # Use ggplot to visualize the matrix
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_tile() +                                       # Better visualization than geom_raster()
  
  # Choose a better color scale
  scale_fill_viridis_c(guide = "none") +

  # Improve axis labels
  scale_x_continuous(
    breaks = seq(0, 200, 50),
    minor_breaks = NULL
  ) +
  scale_y_continuous(
    breaks = seq(0, 200, 50),
    minor_breaks = NULL
  ) +

  # Improve labels and theme
  labs(
    title = "Self-Similarity Matrix of My Second Track",
    x = "Time (s)",
    y = NULL,
    fill = "Distance"
  ) +
  theme_minimal(base_size = 14) +  
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    panel.grid.major = element_line(color = "gray80", size = 0.3),
    panel.grid.minor = element_blank()
  )    
```

### Timbre-based self-similarity for my first track

```{r, fig.width=14, fig.height=7}
similarity_matrix <- "features/aleksandra-b-1.json" |>  
  compmus_mfccs(norm = "identity") |>  
  compmus_self_similarity(
    feature = mfcc,
    distance = "euclidean"
  )

# Debugging: Check if 'd' exists
print(names(similarity_matrix))  # Ensure 'd' is in the data

# Now plot using the stored object
ggplot(similarity_matrix, aes(x = xtime, y = ytime, fill = d)) +  
  geom_tile() +  # Use geom_tile() for better visualization
  
  # Improved color contrast
  scale_fill_viridis_c(guide = "none") +

  # Improved axis labels and readability
  scale_x_continuous(
    breaks = seq(0, 200, 50),
    minor_breaks = NULL
  ) +
  scale_y_continuous(
    breaks = seq(0, 200, 50),
    minor_breaks = NULL
  ) +

  # Add meaningful labels
  labs(
    title = "Timbre-based self-similarity for my first track",
    x = "Time (s)",
    y = NULL,
    fill = "Distance"
  ) +
  theme_minimal(base_size = 14) +  
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    panel.grid.major = element_line(color = "gray80", size = 0.3),
    panel.grid.minor = element_blank()
  )
```

### Timbre-based self-similarity for my second track

```{r, fig.width=14, fig.height=7}
similarity_matrix <- "features/aleksandra-b-2.json" |>  
  compmus_mfccs(norm = "identity") |>  
  compmus_self_similarity(
    feature = mfcc,
    distance = "euclidean"
  )

# Debugging: Check if 'd' exists
print(names(similarity_matrix))  # Ensure 'd' is in the data

# Now plot using the stored object
ggplot(similarity_matrix, aes(x = xtime, y = ytime, fill = d)) +  
  geom_tile() +  # Use geom_tile() for better visualization
  
  # Improved color contrast
  scale_fill_viridis_c(guide = "none") +

  # Improved axis labels and readability
  scale_x_continuous(
    breaks = seq(0, 200, 50),
    minor_breaks = NULL
  ) +
  scale_y_continuous(
    breaks = seq(0, 200, 50),
    minor_breaks = NULL
  ) +

  # Add meaningful labels
  labs(
    title = "Timbre-based self-similarity for my second track",
    x = "Time (s)",
    y = NULL,
    fill = "Distance"
  ) +
  theme_minimal(base_size = 14) +  
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    panel.grid.major = element_line(color = "gray80", size = 0.3),
    panel.grid.minor = element_blank()
  )
```

Chordogram and Keygram {.storyboard}
=========================================
### Chordagram

```{r}

chrdgrm1 <- "features/aleksandra-b-1.json" |> 
  compmus_chroma(norm = "manhattan") |> 
  compmus_match_pitch_templates(
    chord_templates,         # Change to chord_templates if desired
    norm = "manhattan",     # Try different norms (and match it with what you used in `compmus_chroma`)
    distance = "manhattan"   # Try different distance metrics
  ) |>
  ggplot(aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = "Template", fill = NULL) +
  theme_classic()                                      # Change the theme?

chrdgrm1

```

### Keygram

```{r}
chrdgrm2 <- "features/aleksandra-b-2.json" |> 
  compmus_chroma(norm = "euclidean") |> 
  compmus_match_pitch_templates(
    key_templates,         # Change to chord_templates if desired
    norm = "euclidean",      # Try different norms (and match it with what you used in `compmus_chroma`)
    distance = "angular"   # Try different distance metrics
  ) |>
  ggplot(aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = "Template", fill = NULL) +
  theme_classic()                                      # Change the theme?

chrdgrm2
```


Tempograms {.storyboard}
=========================================

### Tempogram for the first track (cyclic=FALSE)

```{r}
"features/aleksandra-b-1.json" |>
  compmus_tempogram(window_size = 8, hop_size = 1, cyclic = FALSE) |>
  ggplot(aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()

```
------------------------------------------------------------------------
The tempograms for my track give some interesting results. The song is an electronic track I fpund on SoundCloud, so I do not know the original data behind it uch as tempo or other elemetns. However, due to the genre of the song, I expected it to have a steady tempo. The first graph was generated with `cyclic = FALSE`, which means it tries to capture all tempo variations separately. This leads to multiple horizontal lines, likely caused by rhythmic layers or subdivisions in the beat. Since the track has strong percussive elements and a lot of repetitive structures, these extra tempo lines are picking up on drum patterns or other rhythmic elements overlapping with the main beat. 

### Tempogram for the first track (cyclic=TRUE)

```{r}
"features/aleksandra-b-1.json" |>
  compmus_tempogram(window_size = 8, hop_size = 1, cyclic = TRUE) |>
  ggplot(aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()
```
------------------------------------------------------------------------
The second graph, made with `cyclic = TRUE`, looks much more stable. Instead of showing multiple tempo components, it focuses on the main beat of the track, which sits clearly around 140 BPM. This makes sense since many electronic dance tracks are designed to have a steady pulse for mixing and dancing. 

In overall, While the first graph captures more rhythmic complexity, the second one highlights the main tempo more clearly.

### Tempogram for the second track (cyclic =FALSE)
```{r}
"features/aleksandra-b-2.json" |>
  compmus_tempogram(window_size = 8, hop_size = 1, cyclic = FALSE) |>
  ggplot(aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()
```
------------------------------------------------------------------------
The next two graphs show the tempo analysis of a track generated by an AI. Even though it was produced not by a human, the tempogram reveals interesting details about how the rhythm is structured. As in the previous tab, the first graph was generated using `cyclic = FALSE`, while the second one was created with `cyclic = TRUE`. Speaking about the first graph, the track has several rhythmic layers or subdivisions, possibly due to programmed beats or overlapping time signatures. The noticeable disruptions in the middle of the graph indicates tempo shifts or sudden changes in rhythmic density, when AI tried to make a transition.

The second graph looks more structured, showing a dominant tempo around 100 BPM with some fluctuations. This suggests that despite the track's complexity, there is a stable underlying beat. The sudden dip in tempo around the middle of the graph indicates a programmed slowdown. Unlike human-composed music, where tempo variations often come from expressive timing, these shifts in an AI-generated track might be a result of algorithmic decisions or transitions between different generated patterns. The first graph highlights the intricate rhythmic elements, while the second one focuses on the main tempo, showing how AI-generated music balances complexity with an overall structured beat.

### Tempogram for the second track (cyclic =TRUE)
```{r}
"features/aleksandra-b-2.json" |>
  compmus_tempogram(window_size = 8, hop_size = 1, cyclic = TRUE) |>
  ggplot(aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()
```
------------------------------------------------------------------------

The second graph looks more structured, showing a dominant tempo around 100 BPM with some fluctuations. This suggests that despite the track's complexity, there is a stable underlying beat. The sudden dip in tempo around the middle of the graph indicates a programmed slowdown. Unlike human-composed music, where tempo variations often come from expressive timing, these shifts in an AI-generated track might be a result of algorithmic decisions or transitions between different generated patterns. The first graph highlights the intricate rhythmic elements, while the second one focuses on the main tempo, showing how AI-generated music balances complexity with an overall structured beat.

### Novelty Functions: Spectral novelty for the 1st track

```{r}
"features/aleksandra-b-1.json" |>
  compmus_spectral_novelty() |> 
  ggplot(aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Time (s)", y = "Spectral Novelty")
```

### Novelty Functions: Spectral novelty for the 2nd track

```{r}
"features/aleksandra-b-2.json" |>
  compmus_spectral_novelty() |> 
  ggplot(aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Time (s)", y = "Spectral Novelty")
```

### Novelty Functions: Energy novelty for the 1st track

```{r}
"features/aleksandra-b-1.json" |>
  compmus_energy_novelty() |> 
  ggplot(aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Time (s)", y = "Energy Novelty")
```

### Novelty Functions: Energy novelty for the 2nd track

```{r}
"features/aleksandra-b-2.json" |>
  compmus_energy_novelty() |> 
  ggplot(aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Time (s)", y = "Energy Novelty")
```

Conclusion {.storyboard}
=========================================
### Conclusion

Songs with a higher tempo (BPM) tend to have a higher arousal level.
Some slower songs still have high energy, possibly due to instrumentation.
Danceability varies across tempos, but higher instrumentalness often lowers danceability.
