---
title: "Music Analysis Dashboard"
author: Alex
output: 
  flexdashboard::flex_dashboard:
    theme: 
      version: 4
      bootswatch: litera
---

```{r setup, include=FALSE}
library(tidyverse)
library(flexdashboard)
library(ggplot2)
library(plotly)
library(rjson)
library(RColorBrewer)
source("compmus.R")

major_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    0,    0)
minor_chord <-
  c(   1,    0,    0,    1,    0,    0,    0,    1,    0,    0,    0,    0)
seventh_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    1,    0)

major_key <-
  c(5.0, 2.0, 3.5, 2.0, 4.5, 4.0, 2.0, 4.5, 2.0, 3.5, 1.5, 4.0)
minor_key <-
  c(5.0, 2.0, 3.5, 4.5, 2.0, 4.0, 2.0, 4.5, 3.5, 2.0, 1.5, 4.0)

chord_templates <-
  tribble(
    ~name, ~template,
    "Gb:7", circshift(seventh_chord, 6),
    "Gb:maj", circshift(major_chord, 6),
    "Bb:min", circshift(minor_chord, 10),
    "Db:maj", circshift(major_chord, 1),
    "F:min", circshift(minor_chord, 5),
    "Ab:7", circshift(seventh_chord, 8),
    "Ab:maj", circshift(major_chord, 8),
    "C:min", circshift(minor_chord, 0),
    "Eb:7", circshift(seventh_chord, 3),
    "Eb:maj", circshift(major_chord, 3),
    "G:min", circshift(minor_chord, 7),
    "Bb:7", circshift(seventh_chord, 10),
    "Bb:maj", circshift(major_chord, 10),
    "D:min", circshift(minor_chord, 2),
    "F:7", circshift(seventh_chord, 5),
    "F:maj", circshift(major_chord, 5),
    "A:min", circshift(minor_chord, 9),
    "C:7", circshift(seventh_chord, 0),
    "C:maj", circshift(major_chord, 0),
    "E:min", circshift(minor_chord, 4),
    "G:7", circshift(seventh_chord, 7),
    "G:maj", circshift(major_chord, 7),
    "B:min", circshift(minor_chord, 11),
    "D:7", circshift(seventh_chord, 2),
    "D:maj", circshift(major_chord, 2),
    "F#:min", circshift(minor_chord, 6),
    "A:7", circshift(seventh_chord, 9),
    "A:maj", circshift(major_chord, 9),
    "C#:min", circshift(minor_chord, 1),
    "E:7", circshift(seventh_chord, 4),
    "E:maj", circshift(major_chord, 4),
    "G#:min", circshift(minor_chord, 8),
    "B:7", circshift(seventh_chord, 11),
    "B:maj", circshift(major_chord, 11),
    "D#:min", circshift(minor_chord, 3)
  )

key_templates <-
  tribble(
    ~name, ~template,
    "Gb:maj", circshift(major_key, 6),
    "Bb:min", circshift(minor_key, 10),
    "Db:maj", circshift(major_key, 1),
    "F:min", circshift(minor_key, 5),
    "Ab:maj", circshift(major_key, 8),
    "C:min", circshift(minor_key, 0),
    "Eb:maj", circshift(major_key, 3),
    "G:min", circshift(minor_key, 7),
    "Bb:maj", circshift(major_key, 10),
    "D:min", circshift(minor_key, 2),
    "F:maj", circshift(major_key, 5),
    "A:min", circshift(minor_key, 9),
    "C:maj", circshift(major_key, 0),
    "E:min", circshift(minor_key, 4),
    "G:maj", circshift(major_key, 7),
    "B:min", circshift(minor_key, 11),
    "D:maj", circshift(major_key, 2),
    "F#:min", circshift(minor_key, 6),
    "A:maj", circshift(major_key, 9),
    "C#:min", circshift(minor_key, 1),
    "E:maj", circshift(major_key, 4),
    "G#:min", circshift(minor_key, 8),
    "B:maj", circshift(major_key, 11),
    "D#:min", circshift(minor_key, 3)
  )
```
Introduction {.storyboard}
=========================================
### Introducing my own tracks

I will structure my portfolio this week, but for now please go to the tempogram section in the menu. Thanks!

### Here is my class corpus visualisation

```{r}
compmus2025 <- read_csv("C:\\Users\\aneru\\musika2\\compmus2025.csv")
compmus2025 |>                   # Start with the data
  ggplot(                     # Set up the plot.
    aes(
      x = valence,
      y = arousal,
      size = tempo,
      colour = danceability
    )
  ) +
  geom_point() +  # Scatter plot.
  geom_smooth() +
  geom_rug(linewidth = 0.1) + # Add 'fringes' to show data distribution.
    scale_x_continuous(         # Fine-tune the x axis.
    limits = c(3, 6.5),
    breaks = c(3, 4, 5, 6, 7), # Specify grid lines
    minor_breaks = NULL       # Remove 'minor' grid lines.
  ) +
  scale_y_continuous(         # Fine-tune the y axis in the same way.
    limits = c(3, 8),
    breaks = c(3, 4, 5, 6, 7, 8),
    minor_breaks = NULL
  ) +
  scale_colour_viridis_c() +  # Use the popular viridis colour palette.
  scale_size_continuous(      # Fine-tune the sizes of each point.
    trans = "exp",            # Use an exp transformation to emphasise loud..
    guide = "none"            # Remove the legend for size.
  ) +
  theme_light() +             # Use a simpler theme.
  labs(                       # Make the titles nice.
    x = "Valence",
    y = "Arousal",
    colour = "Danceability"
  )
ggplotly()
```
------------------------------------------------------------------------

This graph shows the relationship between Valence (how happy or positive a song sounds) and Arousal (how energetic a song is). The x-axis represents Valence, and the y-axis represents Arousal.

Each dot on the graph represents a song. The color of the dots shows Danceability, with purple meaning less danceable and*yellow meaning more danceable. Some dots are bigger than others, which could mean those songs are more popular or appear more often.

There is a blue trend line in the graph, showing that as Valence increases (songs sound happier), Arousal also tends to increase (songs have more energy). The gray area around the line shows some uncertainty, meaning not every song follows this trend exactly.

On the sides of the graph, there are small tick marks (rug plots) that show how many songs fall into different Valence and Arousal levels. 

Overall, the graph suggests that happier songs tend to be more energetic, and danceability varies, with some of the most danceable songs appearing in the high-energy, happy range.

Chromagram {.storyboard}
=========================================

### Chromagram and cepstogram of my first track

```{r fig.show="hold", out.width="49%", fig.align="center"}
"features/aleksandra-b-1.json" |>                       # Changed the track to mine
  compmus_chroma(norm = "manhattan") |>                 # Change the norm
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
   scale_x_continuous(
    breaks = seq(0, 200, 50),  # More structured x-axis ticks
    minor_breaks = NULL
  ) +
  scale_fill_viridis_c(option = "inferno", begin = 0.2, end = 0.8) +  # More readable color scale
  labs(
    title = "Chromagram of my first track",
    x = "Time (s)",
    y = "Pitch Class",
    fill = "Intensity"
  ) +
  theme_minimal(base_size = 14) +  
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    panel.grid.major = element_line(color = "gray80", size = 0.3),
    panel.grid.minor = element_blank()
  )

"features/aleksandra-b-1.json" |>                       # Track file
  compmus_mfccs(norm = "identity") |>                  # MFCC computation
  ggplot(aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = seq(0, 12, 2),  # More structured y-axis
    minor_breaks = NULL
  ) +
  scale_x_continuous(
    breaks = seq(0, 200, 50), # More structured x-axis
    minor_breaks = NULL
  ) +
  scale_fill_distiller(palette = "Spectral", direction = 1) + # Better color contrast
  labs(
    title = "Cepstrogram of my first track",
    x = "Time (s)",
    y = "Coefficient Number",
    fill = "Intensity"
  ) +
  theme_minimal(base_size = 14) +  
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    panel.grid.major = element_line(color = "gray80", size = 0.3),
    panel.grid.minor = element_blank()
  )                              
```
------------------------------------------------------------------------

Text 


### Chromagram and cepstogram of my second track

```{r fig.show="hold", out.width="49%", fig.align="center"}
"features/aleksandra-b-2.json" |>                       # Changed the track to mine
  compmus_chroma(norm = "manhattan") |>                 # Change the norm
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
   scale_x_continuous(
    breaks = seq(0, 200, 50),  # More structured x-axis ticks
    minor_breaks = NULL
  ) +
  scale_fill_viridis_c(option = "cividis", begin = 0.2, end = 0.8) +  # More readable color scale
  labs(
    title = "Chromagram of my second track",
    x = "Time (s)",
    y = "Pitch Class",
    fill = "Intensity"
  ) +
  theme_minimal(base_size = 14) +  
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    panel.grid.major = element_line(color = "gray80", size = 0.3),
    panel.grid.minor = element_blank()
  )

"features/aleksandra-b-2.json" |>                       # Track file
  compmus_mfccs(norm = "identity") |>                  # MFCC computation
  ggplot(aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = seq(0, 12, 2),  # More structured y-axis
    minor_breaks = NULL
  ) +
  scale_x_continuous(
    breaks = seq(0, 200, 50), # More structured x-axis
    minor_breaks = NULL
  ) +
  scale_fill_distiller(palette = "RdYlBu", direction = 1) + # Better color contrast
  labs(
    title = "Cepstrogram of my second track",
    x = "Time (s)",
    y = "Coefficient Number",
    fill = "Intensity"
  ) +
  theme_minimal(base_size = 14) +  
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    panel.grid.major = element_line(color = "gray80", size = 0.3),
    panel.grid.minor = element_blank()
  )                              
```

------------------------------------------------------------------------
Text 



Self-Similarity Matrices {.storyboard}
=========================================
### Chroma-based self-similarity for my first track

```{r, fig.width=14, fig.height=7}
"features/aleksandra-b-1.json" |>                     # Change the track
  compmus_chroma(norm = "identity") |>                # Change the norm
  compmus_self_similarity(
    feature = pc,
    distance = "euclidean"                            # Change the distance metric
  ) |>
  
  # Use ggplot to visualize the matrix
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_tile() +                                       # Better visualization than geom_raster()
  
  # Choose a better color scale
  scale_fill_viridis_c(guide = "none") +

  # Improve axis labels
  scale_x_continuous(
    breaks = seq(0, 200, 50),
    minor_breaks = NULL
  ) +
  scale_y_continuous(
    breaks = seq(0, 200, 50),
    minor_breaks = NULL
  ) +

  # Improve labels and theme
  labs(
    title = "Self-Similarity Matrix of My First Track",
    x = "Time (s)",
    y = NULL,
    fill = "Distance"
  ) +
  theme_minimal(base_size = 14) +  
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    panel.grid.major = element_line(color = "gray80", size = 0.3),
    panel.grid.minor = element_blank()
  )    
```

### Chroma-based self-similarity for my second track

```{r, fig.width=14, fig.height=7}
"features/aleksandra-b-2.json" |>                     # Change the track
  compmus_chroma(norm = "identity") |>                # Change the norm
  compmus_self_similarity(
    feature = pc,
    distance = "euclidean"                            # Change the distance metric
  ) |>
  
  # Use ggplot to visualize the matrix
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_tile() +                                       # Better visualization than geom_raster()
  
  # Choose a better color scale
  scale_fill_viridis_c(guide = "none") +

  # Improve axis labels
  scale_x_continuous(
    breaks = seq(0, 200, 50),
    minor_breaks = NULL
  ) +
  scale_y_continuous(
    breaks = seq(0, 200, 50),
    minor_breaks = NULL
  ) +

  # Improve labels and theme
  labs(
    title = "Self-Similarity Matrix of My Second Track",
    x = "Time (s)",
    y = NULL,
    fill = "Distance"
  ) +
  theme_minimal(base_size = 14) +  
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    panel.grid.major = element_line(color = "gray80", size = 0.3),
    panel.grid.minor = element_blank()
  )    
```

### Timbre-based self-similarity for my first track

```{r, fig.width=14, fig.height=7}
similarity_matrix <- "features/aleksandra-b-1.json" |>  
  compmus_mfccs(norm = "identity") |>  
  compmus_self_similarity(
    feature = mfcc,
    distance = "euclidean"
  )

# Debugging: Check if 'd' exists
print(names(similarity_matrix))  # Ensure 'd' is in the data

# Now plot using the stored object
ggplot(similarity_matrix, aes(x = xtime, y = ytime, fill = d)) +  
  geom_tile() +  # Use geom_tile() for better visualization
  
  # Improved color contrast
  scale_fill_viridis_c(guide = "none") +

  # Improved axis labels and readability
  scale_x_continuous(
    breaks = seq(0, 200, 50),
    minor_breaks = NULL
  ) +
  scale_y_continuous(
    breaks = seq(0, 200, 50),
    minor_breaks = NULL
  ) +

  # Add meaningful labels
  labs(
    title = "Timbre-based self-similarity for my first track",
    x = "Time (s)",
    y = NULL,
    fill = "Distance"
  ) +
  theme_minimal(base_size = 14) +  
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    panel.grid.major = element_line(color = "gray80", size = 0.3),
    panel.grid.minor = element_blank()
  )
```

### Timbre-based self-similarity for my second track

```{r, fig.width=14, fig.height=7}
similarity_matrix <- "features/aleksandra-b-2.json" |>  
  compmus_mfccs(norm = "identity") |>  
  compmus_self_similarity(
    feature = mfcc,
    distance = "euclidean"
  )

# Debugging: Check if 'd' exists
print(names(similarity_matrix))  # Ensure 'd' is in the data

# Now plot using the stored object
ggplot(similarity_matrix, aes(x = xtime, y = ytime, fill = d)) +  
  geom_tile() +  # Use geom_tile() for better visualization
  
  # Improved color contrast
  scale_fill_viridis_c(guide = "none") +

  # Improved axis labels and readability
  scale_x_continuous(
    breaks = seq(0, 200, 50),
    minor_breaks = NULL
  ) +
  scale_y_continuous(
    breaks = seq(0, 200, 50),
    minor_breaks = NULL
  ) +

  # Add meaningful labels
  labs(
    title = "Timbre-based self-similarity for my second track",
    x = "Time (s)",
    y = NULL,
    fill = "Distance"
  ) +
  theme_minimal(base_size = 14) +  
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    panel.grid.major = element_line(color = "gray80", size = 0.3),
    panel.grid.minor = element_blank()
  )
```

Chordogram and Keygram {.storyboard}
=========================================
### Chordagrams

```{r, out.width="100%", out.height="60%"}

chrdgrm1 <- "features/aleksandra-b-1.json" |> 
  compmus_chroma(norm = "manhattan") |> 
  compmus_match_pitch_templates(
    chord_templates,         # Change to chord_templates if desired
    norm = "manhattan",     # Try different norms (and match it with what you used in `compmus_chroma`)
    distance = "manhattan"   # Try different distance metrics
  ) |>
  ggplot(aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = "Template", fill = NULL) +
  theme_classic()                                      # Change the theme?

chrdgrm1

chrdgrm2 <- "features/aleksandra-b-2.json" |> 
  compmus_chroma(norm = "euclidean") |> 
  compmus_match_pitch_templates(
    key_templates,         # Change to chord_templates if desired
    norm = "euclidean",      # Try different norms (and match it with what you used in `compmus_chroma`)
    distance = "angular"   # Try different distance metrics
  ) |>
  ggplot(aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = "Template", fill = NULL) +
  theme_classic()                                      # Change the theme?

chrdgrm2

```

Tempograms {.storyboard}
=========================================

### Tempograms for the first track

```{r fig.show="hold", out.width="49%", fig.align="center"}
"features/aleksandra-b-1.json" |>
  compmus_tempogram(window_size = 8, hop_size = 1, cyclic = FALSE) |>
  ggplot(aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()

"features/aleksandra-b-1.json" |>
  compmus_tempogram(window_size = 8, hop_size = 1, cyclic = TRUE) |>
  ggplot(aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()
```


### Tempograms for the second track
```{r fig.show="hold", out.width="49%", fig.align="center"}
"features/aleksandra-b-2.json" |>
  compmus_tempogram(window_size = 8, hop_size = 1, cyclic = FALSE) |>
  ggplot(aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()

"features/aleksandra-b-2.json" |>
  compmus_tempogram(window_size = 8, hop_size = 1, cyclic = TRUE) |>
  ggplot(aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()
```

### Novelty Functions: Spectral novelty
```{r fig.show="hold", out.width="49%", fig.align="center"}
"features/aleksandra-b-1.json" |>
  compmus_spectral_novelty() |> 
  ggplot(aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Time (s)", y = "Spectral Novelty")

"features/aleksandra-b-2.json" |>
  compmus_spectral_novelty() |> 
  ggplot(aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Time (s)", y = "Spectral Novelty")
```


### Novelty Functions: Energy novelty
```{r fig.show="hold", out.width="49%", fig.align="center"}
"features/aleksandra-b-1.json" |>
  compmus_energy_novelty() |> 
  ggplot(aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Time (s)", y = "Energy Novelty")

"features/aleksandra-b-2.json" |>
  compmus_energy_novelty() |> 
  ggplot(aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Time (s)", y = "Energy Novelty")
```

Conclusion {.storyboard}
=========================================
### Conclusion

Songs with a higher tempo (BPM) tend to have a higher arousal level.
Some slower songs still have high energy, possibly due to instrumentation.
Danceability varies across tempos, but higher instrumentalness often lowers danceability.
