---
title: "Music Analysis Dashboard"
author: Alex
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
---
```{r}
library(tidyverse)
source("compmus.R")
```

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)  
library(plotly)
library(ggrepel)  
library(viridis)

```
-----------------------------------------------------------------------
### Chromagram of my first track

```{r, fig.width=12, fig.height=6}
"features/aleksandra-b-1.json" |>                       # Changed the track to mine
  compmus_chroma(norm = "manhattan") |>                 # Change the norm
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
   scale_x_continuous(
    breaks = seq(0, 200, 50),  # More structured x-axis ticks
    minor_breaks = NULL
  ) +
  scale_fill_viridis_c(option = "inferno", begin = 0.2, end = 0.8) +  # More readable color scale
  labs(
    title = "Chromagram of my first track",
    x = "Time (s)",
    y = "Pitch Class",
    fill = "Intensity"
  ) +
  theme_minimal(base_size = 14) +  
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    panel.grid.major = element_line(color = "gray80", size = 0.3),
    panel.grid.minor = element_blank()
  )
```

### Chromagram of my second track
```{r, fig.width=12, fig.height=6}
"features/aleksandra-b-2.json" |>                       # Changed the track to mine
  compmus_chroma(norm = "manhattan") |>                 # Change the norm
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
   scale_x_continuous(
    breaks = seq(0, 200, 50),  # More structured x-axis ticks
    minor_breaks = NULL
  ) +
  scale_fill_viridis_c(option = "cividis", begin = 0.2, end = 0.8) +  # More readable color scale
  labs(
    title = "Chromagram of my second track",
    x = "Time (s)",
    y = "Pitch Class",
    fill = "Intensity"
  ) +
  theme_minimal(base_size = 14) +  
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    panel.grid.major = element_line(color = "gray80", size = 0.3),
    panel.grid.minor = element_blank()
  )
```



### Cepstrogram of my first track

```{r}
library(RColorBrewer)
```

```{r, fig.width=14, fig.height=6}
"features/aleksandra-b-1.json" |>                       # Track file
  compmus_mfccs(norm = "identity") |>                  # MFCC computation
  ggplot(aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = seq(0, 12, 2),  # More structured y-axis
    minor_breaks = NULL
  ) +
  scale_x_continuous(
    breaks = seq(0, 200, 50), # More structured x-axis
    minor_breaks = NULL
  ) +
  scale_fill_distiller(palette = "Spectral", direction = 1) + # Better color contrast
  labs(
    title = "Cepstrogram of my first track",
    x = "Time (s)",
    y = "Coefficient Number",
    fill = "Intensity"
  ) +
  theme_minimal(base_size = 14) +  
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    panel.grid.major = element_line(color = "gray80", size = 0.3),
    panel.grid.minor = element_blank()
  )                              
```

### Cepstrogram of my second track

```{r, fig.width=14, fig.height=6}
"features/aleksandra-b-2.json" |>                       # Track file
  compmus_mfccs(norm = "identity") |>                  # MFCC computation
  ggplot(aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = seq(0, 12, 2),  # More structured y-axis
    minor_breaks = NULL
  ) +
  scale_x_continuous(
    breaks = seq(0, 200, 50), # More structured x-axis
    minor_breaks = NULL
  ) +
  scale_fill_distiller(palette = "RdYlBu", direction = 1) + # Better color contrast
  labs(
    title = "Cepstrogram of my second track",
    x = "Time (s)",
    y = "Coefficient Number",
    fill = "Intensity"
  ) +
  theme_minimal(base_size = 14) +  
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    panel.grid.major = element_line(color = "gray80", size = 0.3),
    panel.grid.minor = element_blank()
  )                              
```

### Chroma-based self-similarity for my first track

```{r, fig.width=14, fig.height=7}
"features/aleksandra-b-1.json" |>                     # Change the track
  compmus_chroma(norm = "identity") |>                # Change the norm
  compmus_self_similarity(
    feature = pc,
    distance = "euclidean"                            # Change the distance metric
  ) |>
  
  # Use ggplot to visualize the matrix
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_tile() +                                       # Better visualization than geom_raster()
  
  # Choose a better color scale
  scale_fill_viridis_c(guide = "none") +

  # Improve axis labels
  scale_x_continuous(
    breaks = seq(0, 200, 50),
    minor_breaks = NULL
  ) +
  scale_y_continuous(
    breaks = seq(0, 200, 50),
    minor_breaks = NULL
  ) +

  # Improve labels and theme
  labs(
    title = "Self-Similarity Matrix of My First Track",
    x = "Time (s)",
    y = NULL,
    fill = "Distance"
  ) +
  theme_minimal(base_size = 14) +  
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    panel.grid.major = element_line(color = "gray80", size = 0.3),
    panel.grid.minor = element_blank()
  )    
```

### Chroma-based self-similarity for my second track

```{r, fig.width=14, fig.height=7}
"features/aleksandra-b-2.json" |>                     # Change the track
  compmus_chroma(norm = "identity") |>                # Change the norm
  compmus_self_similarity(
    feature = pc,
    distance = "euclidean"                            # Change the distance metric
  ) |>
  
  # Use ggplot to visualize the matrix
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_tile() +                                       # Better visualization than geom_raster()
  
  # Choose a better color scale
  scale_fill_viridis_c(guide = "none") +

  # Improve axis labels
  scale_x_continuous(
    breaks = seq(0, 200, 50),
    minor_breaks = NULL
  ) +
  scale_y_continuous(
    breaks = seq(0, 200, 50),
    minor_breaks = NULL
  ) +

  # Improve labels and theme
  labs(
    title = "Self-Similarity Matrix of My Second Track",
    x = "Time (s)",
    y = NULL,
    fill = "Distance"
  ) +
  theme_minimal(base_size = 14) +  
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    panel.grid.major = element_line(color = "gray80", size = 0.3),
    panel.grid.minor = element_blank()
  )    
```

### Timbre-based self-similarity for my first track

```{r, fig.width=14, fig.height=7}
similarity_matrix <- "features/aleksandra-b-1.json" |>  
  compmus_mfccs(norm = "identity") |>  
  compmus_self_similarity(
    feature = mfcc,
    distance = "euclidean"
  )

# Debugging: Check if 'd' exists
print(names(similarity_matrix))  # Ensure 'd' is in the data

# Now plot using the stored object
ggplot(similarity_matrix, aes(x = xtime, y = ytime, fill = d)) +  
  geom_tile() +  # Use geom_tile() for better visualization
  
  # Improved color contrast
  scale_fill_viridis_c(guide = "none") +

  # Improved axis labels and readability
  scale_x_continuous(
    breaks = seq(0, 200, 50),
    minor_breaks = NULL
  ) +
  scale_y_continuous(
    breaks = seq(0, 200, 50),
    minor_breaks = NULL
  ) +

  # Add meaningful labels
  labs(
    title = "Timbre-based self-similarity for my first track",
    x = "Time (s)",
    y = NULL,
    fill = "Distance"
  ) +
  theme_minimal(base_size = 14) +  
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    panel.grid.major = element_line(color = "gray80", size = 0.3),
    panel.grid.minor = element_blank()
  )
```

### Timbre-based self-similarity for my second track

```{r, fig.width=14, fig.height=7}
similarity_matrix <- "features/aleksandra-b-2.json" |>  
  compmus_mfccs(norm = "identity") |>  
  compmus_self_similarity(
    feature = mfcc,
    distance = "euclidean"
  )

# Debugging: Check if 'd' exists
print(names(similarity_matrix))  # Ensure 'd' is in the data

# Now plot using the stored object
ggplot(similarity_matrix, aes(x = xtime, y = ytime, fill = d)) +  
  geom_tile() +  # Use geom_tile() for better visualization
  
  # Improved color contrast
  scale_fill_viridis_c(guide = "none") +

  # Improved axis labels and readability
  scale_x_continuous(
    breaks = seq(0, 200, 50),
    minor_breaks = NULL
  ) +
  scale_y_continuous(
    breaks = seq(0, 200, 50),
    minor_breaks = NULL
  ) +

  # Add meaningful labels
  labs(
    title = "Timbre-based self-similarity for my second track",
    x = "Time (s)",
    y = NULL,
    fill = "Distance"
  ) +
  theme_minimal(base_size = 14) +  
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    panel.grid.major = element_line(color = "gray80", size = 0.3),
    panel.grid.minor = element_blank()
  )
```

### Introduction: Understanding the Dataset (First Homework)
Introduction
This dashboard explores a dataset of songs, analyzing their tempo, arousal, and danceability.
We will visualize key characteristics and provide insights on how they relate to each other.

Dataset Overview
Arousal → The energy level of a song.
Tempo (BPM) → The speed of the song.
Danceability → How easy it is to dance to the song.
Instrumentalness → How much of the track is instrumental.


### Visualisation: Tempo vs Arousal
```{r}
# Load dataset
aisc2024 <- read_csv("aisc2024.csv")

# Ensure dataset exists
if (!exists("aisc2024") || nrow(aisc2024) == 0) {
  stop("Error: The dataset 'aisc2024' is not found or empty.")
}

# Create base ggplot
p <- ggplot(aisc2024, aes(
    x = tempo,
    y = arousal,
    size = instrumentalness,
    colour = danceability
  )) +
  geom_point(alpha = 0.8) +  
  geom_rug(linewidth = 0.2, sides = "b") +  
  scale_x_continuous(limits = c(50, 200), breaks = seq(50, 200, 50)) +
  scale_y_continuous(limits = c(1, 9), breaks = c(1, 5, 9)) +
  scale_colour_viridis_c(option = "C") +  
  scale_size_continuous(trans = "sqrt", guide = "none") +
  theme_minimal(base_size = 14) +  
  theme(
    legend.position = "right",
    axis.title = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    panel.grid.major = element_line(size = 0.3, linetype = "dashed"),
    panel.grid.minor = element_blank()
  ) +
  labs(
    title = "Music Characteristics: Tempo vs Arousal",
    x = "Tempo (BPM)",
    y = "Arousal Level",
    colour = "Danceability"
  )

# Static ggplot2 (with text labels)
p_static <- p + 
  geom_text_repel(
    aes(label = ifelse(tempo > 120 & arousal > 4, "Onda Corta - Sud America", NA)),  
    size = 4,
    fontface = "bold",
    color = "black",
    nudge_x = 10, 
    nudge_y = 0.5,
    box.padding = 0.4,
    point.padding = 0.4,
    segment.color = "gray50",
    max.overlaps = 3  
  )

# Interactive ggplotly (without text labels)
p_interactive <- ggplotly(p)

# Render the correct plot
p_interactive  # This ensures the plot is displayed in Flexdashboard


```

### Conclusion
Songs with a higher tempo (BPM) tend to have a higher arousal level.
Some slower songs still have high energy, possibly due to instrumentation.
Danceability varies across tempos, but higher instrumentalness often lowers danceability.




